---
title: "GEO 880 - week 5 - Tasks and Inputs"
author: "Lars Weidinger"
date: "today"
format: 
  html:
    fig_caption: true
    fig-align: center
    highlight: tango
    number-sections: false
    code-fold: true
    theme: lumen
    embed-resources: true
    toc: true
    toc_depth: 2
    toc_float: true
execute:                      # Set global options for execution of code blocks
  echo: true                  # False = no code printed
  warning: false
  message: false
editor: visual
---

<!-- The following html code fragment produces fully justified text alignment -->

```{=html}
<style>
body {text-align: justify}
</style>
```

# Preliminaries

## Set Options

```{r}
options(warning = FALSE)    # Don't show warnings
```

## Install or Load packages

```{r}
#| code-fold: show
library(pacman)

# Load and if not found in the library will attempt to install the package
pacman::p_load(here,
               sf,
               readr,
               tmap,
               dplyr,
               ggplot2,
               tidyr,
               SimilarityMeasures,
               lubridate,
               plotly,
               install = TRUE)
```

## Store relevant CRS

```{r}
crs_lv03  <- 21781
crs_lv95  <- 2056
crs_wgs84 <- 4326
```

# Task 1: Write your own functions
## 1)
A function which calculates a persons BMI based on their height and weight.
```{r}
BMI <- function(weight, height) {
    BMI_value <- weight / (height^2)
    return(BMI_value)
}

BMI(weight = 78, height = 1.8)
```

## 2)
A function which converts degrees Celsius to Fahrenheit
```{r}
C_to_F <- function(celsius) {
  (celsius * (9/5) + 32)
}

C_to_F(celsius = 20)
```

## 3)
A function which calculates the (Euclidean) distance between two sets of coordinates.
```{r}
fundis <- function(x1, y1, x2, y2) {
  sqrt( ((x2-x1)^2) + ((y2-y1)^2) )
}

fundis(x1 = 10, y1 = 0, x2 = 0, y2 = 0)
```

# Task 2: Prepare Analysis
```{r}
wildschwein <- read_delim("data/wildschwein_BE_2056.csv", ",")

# Careful! What Timezone is assumed? --> UTC
wildschwein <- wildschwein |>
    filter(
      TierName == "Rosa" | TierName == "Sabi", 
      DatetimeUTC >= "2015-04-01", 
      DatetimeUTC < "2015-04-15"
      )
```

# Task 3: Create Join Key
```{r}
wildschwein <- wildschwein |>
  mutate(
    DatetimeRound = lubdridate::round_date(DatetimeUTC, unit = "15 minutes") # so that we can later match them exactely
    )
```

# Task 4: Measuring distance at concurrent locations
## 1) Split
```{r}
rosa <- wildschwein |>
  filter(TierName == "Rosa")

sabi <- wildschwein |>
  filter(TierName == "Sabi")
```

## 2) Join
```{r}
# we can do a full join so that everything is included

wildschwein_join <- full_join(rosa, sabi, by = c("DatetimeRound"), suffix = c("_rosa", "_sabi"))
```

## 3 & 4) Euclidian Distance & Spatially Close
```{r}
wildschwein_join <- wildschwein_join |>
  mutate(
    distance = sqrt( (N_rosa - N_sabi)^2 + (E_rosa - E_sabi)^2 ), # distance between the two fixes that are temporally close
    meet = distance < 100 # should give T/F
  )
```

# Task 5: Visualize data
```{r}
wildschwein_meet <- wildschwein_join |>
  filter(meet)

ggplot(wildschwein_meet) +
  
  geom_point(data = rosa, aes(E, N, colour = "rosa"), alpha = 0.5) + # this is only rosa
  geom_point(data = sabi, aes(E, N, colour = "sabi"), alpha = 0.5) + # this is only sabi
  
  geom_point(aes(x = E_rosa, y = N_rosa, fill = "rosa")) + 
  geom_point(aes(x = E_sabi, y = N_sabi, fill = "sabi")) +
  
  coord_equal() +
  
  theme_minimal()
```

